---
title: "Long spatiotemporal indices of biomass density by benthic species group"
author: "Max Lindmark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  df_print: paged
pdf_document: default
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 12,
  fig.asp = 0.618,
  fig.align ='center'
)
```

## Load libraries
```{r libraries, message=FALSE}
# Load libraries, install if needed
library(tidyverse); theme_set(theme_classic())
library(readxl)
library(tidylog)
library(RCurl)
library(sp)
library(geosphere)
library(viridis)
library(RColorBrewer)
library(patchwork)
library(janitor)
library(icesDatras)
library(mapdata)
library(patchwork)
library(rgdal)
library(raster)
library(sf)
library(rgeos)
library(chron)
library(lattice)
library(ncdf4)
library(sdmTMB) # remotes::install_github("pbs-assess/sdmTMB")
library(marmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(mapplots)
library(mgcv)
library(qwraps2) # To load entire cache in interactive r session, do: qwraps2::lazyload_cache_dir(path = "R/spatio_temporal_indices/long_benthic_index_cache/html")

# For adding maps to plots
world <- ne_countries(scale = "medium", returnclass = "sf")

# Specify map ranges
ymin = 55.5; ymax = 58; xmin = 12.5; xmax = 20
```


# Saduria
Read data

```{r sad data, message=FALSE}
# sad <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/bentfish/main/data/for_analysis/saduria.csv") %>%
#   dplyr::select(-X1)
sad <- read.csv("data/for_analysis/saduria.csv") %>% dplyr::select(-X)

sad <- sad %>%
  filter(year > 1985) %>% 
  mutate(SubDiv = as.factor(SubDiv)) %>% 
  filter(SubDiv %in% c(25, 27, 28)) %>% 
  mutate(year_f = as.factor(year_f),
         depth_scaled_sq = depth_scaled*depth_scaled)

# Split data by subdivison
sad25 <- sad %>% filter(SubDiv == "25")
sad25 %>%
  group_by(year) %>%
  summarize(n = n()) %>%
  ggplot(., aes(year, n)) + geom_bar(stat = "identity")

ggplot(sad25, aes(lon, lat, color = biomass)) + geom_point()
ggplot(sad25, aes(biomass)) + geom_histogram()

sad27 <- sad %>% filter(SubDiv == "27" & lon < 17.4)
sad27 %>%
  group_by(year) %>%
  summarize(n = n()) %>%
  ggplot(., aes(year, n)) + geom_bar(stat = "identity")

ggplot(sad27, aes(lon, lat, color = biomass)) + geom_point()
ggplot(sad27, aes(biomass)) + geom_histogram()

# remove outlier and see if model improves
sad27 <- sad27 %>% filter(biomass < 80)

sad28 <- sad %>% filter(SubDiv == "28")
sad28 %>%
  group_by(year) %>%
  summarize(n = n()) %>%
  ggplot(., aes(year, n)) + geom_bar(stat = "identity")

ggplot(sad28, aes(biomass)) + geom_histogram()

sad28 <- sad28 %>% filter(year > 2005)

# remove data points wiht more than 50 g/m2
sad25 <- sad25 %>% filter(biomass < 80)
sad27 <- sad27 %>% filter(biomass < 80)
sad28 <- sad28 %>% filter(biomass < 80)
```

Read and crop pred grid to match saduria

Fit models using `sdmTMB` assuming a Tweedie distribution and biomass as the response.

Make spde mesh

```{r make sad barrier spde mesh, results='hide', cache=TRUE, message=FALSE}
# Non-island version
sad_spde25 <- make_mesh(data = sad25, xy_cols = c("lon", "lat"), n_knots = 50, type = "kmeans", seed = 42)

sad_spde27 <- make_mesh(data = sad27, xy_cols = c("lon", "lat"), n_knots = 40, type = "kmeans", seed = 42)

sad_spde28 <- make_mesh(data = sad28, xy_cols = c("lon", "lat"), n_knots = 20, type = "kmeans", seed = 42)
```

Fit the models

```{r fit sad tweedie model, cache=TRUE}
m_sad25 <- sdmTMB(formula = biomass ~ year_f - 1 + depth_scaled + depth_scaled_sq, data = sad25,
                  time = "year", spde = sad_spde25, family = tweedie(link = "log"),
                  ar1_fields = FALSE, include_spatial = TRUE, spatial_trend = FALSE, spatial_only = TRUE)

m_sad27 <- sdmTMB(formula = biomass ~ year_f - 1 + depth_scaled + depth_scaled_sq, data = sad27,
                  time = "year", spde = sad_spde27, family = tweedie(link = "log"),
                  ar1_fields = FALSE, include_spatial = TRUE, spatial_trend = FALSE, spatial_only = TRUE)

m_sad28 <- sdmTMB(formula = biomass ~ year_f - 1 + depth_scaled + depth_scaled_sq, data = sad28,
                  time = "year", spde = sad_spde28, family = tweedie(link = "log"),
                  ar1_fields = FALSE, include_spatial = TRUE, spatial_trend = FALSE, spatial_only = TRUE)

# Check residuals
sad25$residuals <- residuals(m_sad25)
sad27$residuals <- residuals(m_sad27)
sad28$residuals <- residuals(m_sad28)

# Pretty good!
qqnorm(sad25$residuals); abline(a = 0, b = 1)
qqnorm(sad27$residuals); abline(a = 0, b = 1)
qqnorm(sad28$residuals); abline(a = 0, b = 1)

# Plot the marginal effect of depth:
# SD 25
nd_sad25 <- data.frame(depth_scaled = seq(min(sad25$depth_scaled), max(sad25$depth_scaled), length.out = 100))
nd_sad25$depth_scaled_sq <- nd_sad25$depth_scaled*nd_sad25$depth_scaled
nd_sad25$year <- as.integer(max(sad25$year))
nd_sad25$year_f <- factor(nd_sad25$year)

p_sad25 <- predict(m_sad25, newdata = nd_sad25, se_fit = TRUE, re_form = NA)

ggplot(p_sad25, aes(depth_scaled, exp(est),
                    ymin = exp(est - 1.96 * est_se), ymax = exp(est + 1.96 * est_se))) +
  geom_line() + geom_ribbon(alpha = 0.4)

# SD 27
nd_sad27 <- data.frame(depth_scaled = seq(min(sad27$depth_scaled), max(sad27$depth_scaled), length.out = 100))
nd_sad27$depth_scaled_sq <- nd_sad27$depth_scaled*nd_sad27$depth_scaled
nd_sad27$year <- as.integer(max(sad27$year))
nd_sad27$year_f <- factor(nd_sad27$year)

p_sad27 <- predict(m_sad27, newdata = nd_sad27, se_fit = TRUE, re_form = NA)

ggplot(p_sad27, aes(depth_scaled, exp(est),
                    ymin = exp(est - 1.96 * est_se), ymax = exp(est + 1.96 * est_se))) +
  geom_line() + geom_ribbon(alpha = 0.4)

# SD 28
nd_sad28 <- data.frame(depth_scaled = seq(min(sad28$depth_scaled), max(sad28$depth_scaled), length.out = 100))
nd_sad28$depth_scaled_sq <- nd_sad28$depth_scaled*nd_sad28$depth_scaled
nd_sad28$year <- as.integer(max(sad28$year))
nd_sad28$year_f <- factor(nd_sad28$year)

p_sad28 <- predict(m_sad28, newdata = nd_sad28, se_fit = TRUE, re_form = NA)

ggplot(p_sad28, aes(depth_scaled, exp(est),
                    ymin = exp(est - 1.96 * est_se), ymax = exp(est + 1.96 * est_se))) +
  geom_line() + geom_ribbon(alpha = 0.4)
```

Read in the prediction grids
Read and crop pred grid to match saduria

```{r crop pred grid for saduria, message=FALSE, warning=FALSE}
## Read the prediction grid
# pred_grid <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/bentfish/main/data/for_analysis/pred_grid.csv")
pred_grid <- read.csv("data/for_analysis/pred_grid.csv")

pred_grid <- pred_grid %>%
  mutate(depth_scaled = (depth - mean(sad$depth))/sd(sad$depth)) %>% 
  mutate(depth_scaled_sq = depth_scaled*depth_scaled) %>% 
  mutate(year_f = factor(year),
         year = as.integer(year))

# Subset the prediction grid to match saduria data
pred_grid_sad25 <- pred_grid %>% filter(SubDiv == "25" & year %in% c(unique(sad25$year)))
pred_grid_sad27 <- pred_grid %>% filter(SubDiv == "27" & year %in% c(unique(sad27$year)))
pred_grid_sad28 <- pred_grid %>% filter(SubDiv == "28" & year %in% c(unique(sad28$year)))

tf_sad25 <- exclude.too.far(pred_grid_sad25$lon, pred_grid_sad25$lat, sad$lon, sad$lat, 0.05) # 0.03 seems reasonable
tf_sad27 <- exclude.too.far(pred_grid_sad27$lon, pred_grid_sad27$lat, sad$lon, sad$lat, 0.05) # 0.03 seems reasonable
tf_sad28 <- exclude.too.far(pred_grid_sad28$lon, pred_grid_sad28$lat, sad$lon, sad$lat, 0.05) # 0.03 seems reasonable

# Filter the grid points that are not too far from the data
pred_grid_sad25$too_far <- tf_sad25
pred_grid_sad27$too_far <- tf_sad27
pred_grid_sad28$too_far <- tf_sad28

# Plot again
pred_grid_sad25 %>%
  filter(too_far == FALSE) %>%
  ggplot(., aes(lon, lat, fill = factor(SubDiv))) +
  geom_raster() +
  geom_point(data = sad25, aes(lon, lat), color = "black", size = 0.5) +
  NULL

pred_grid_sad27 %>%
  filter(too_far == FALSE) %>%
  ggplot(., aes(lon, lat, fill = factor(SubDiv))) +
  geom_raster() +
  geom_point(data = sad27, aes(lon, lat), color = "black", size = 0.5) +
  NULL

pred_grid_sad28 %>%
  filter(too_far == FALSE) %>%
  ggplot(., aes(lon, lat, fill = factor(SubDiv))) +
  geom_raster() +
  geom_point(data = sad28, aes(lon, lat), color = "black", size = 0.5) +
  NULL
```

Now calculate the index by year and sub area

```{r calculate sad index, cache=TRUE, message=FALSE}
# From these models, predict annual biomass for each sub area
# sort(unique(sad$SubDiv))

# Sub division 25
sad_preds25 <- predict(m_sad25, pred_grid_sad25, return_tmb_object = TRUE) # Predict using the grid for subarea

sad_ind25 <- get_index(sad_preds25, bias_correct = T) # Get the index (sum of all cells in the pred grid)

# The index is now the sum of grid cells in g/area of a grid cell
# First multiply by the area of a grid cell to get raw g
ncells25 <- pred_grid_sad25 %>% filter(year == 2015 & SubDiv == 25) %>% nrow()
# Then divide that by the total area

# This is the same as simply dividing by ncells25
sad_ind25 <- sad_ind25 %>% mutate(est = est/ncells25, lwr = lwr/ncells25, upr = upr/ncells25, SubDiv = 25)


# Sub division 27
sad_preds27 <- predict(m_sad27, pred_grid_sad27, return_tmb_object = TRUE) # Predict using the grid for subarea

sad_ind27 <- get_index(sad_preds27, bias_correct = F) # Get the index (sum of all cells in the pred grid)

# The index is now the sum of grid cells in g/area of a grid cell
# First multiply by the area of a grid cell to get raw g
ncells27 <- pred_grid_sad27 %>% filter(year == 2015 & SubDiv == 27) %>% nrow()
# Then divide that by the total area

# This is the same as simply dividing by ncells27
sad_ind27 <- sad_ind27 %>% mutate(est = est/ncells27, lwr = lwr/ncells27, upr = upr/ncells27, SubDiv = 27)


# Sub division 28
sad_preds28 <- predict(m_sad28, pred_grid_sad28, return_tmb_object = TRUE) # Predict using the grid for subarea

sad_ind28 <- get_index(sad_preds28, bias_correct = T) # Get the index (sum of all cells in the pred grid)

# The index is now the sum of grid cells in g/area of a grid cell
# First multiply by the area of a grid cell to get raw g
ncells28 <- pred_grid_sad28 %>% filter(year == 2015 & SubDiv == 28) %>% nrow()
# Then divide that by the total area

# This is the same as simply dividing by ncells28
sad_ind28 <- sad_ind28 %>% mutate(est = est/ncells28, lwr = lwr/ncells28, upr = upr/ncells28, SubDiv = 28)


# Merge prediction-data
sad_preds <- bind_rows(sad_ind25, sad_ind27, sad_ind28) %>% mutate(SubDiv = factor(SubDiv))

# Compare to data quickly
p <- sad_preds %>%
  dplyr::select(year, est, upr, lwr, SubDiv) %>%
  mutate(source = "pred") %>% 
  arrange(year, SubDiv)

d <- sad %>%
  group_by(year, SubDiv) %>% 
  summarise(est = mean(biomass),
            sd = sd(biomass)) %>%
  mutate(source = "data") %>% 
  arrange(year, SubDiv)

ggplot(bind_rows(p, d), aes(year, est, color = source)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = est-sd, ymax = est+sd)) + 
  geom_line(size = 1) + 
  facet_wrap(~ SubDiv, scales = "free", ncol = 4) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.4) + 
  theme_classic(base_size = 15) + 
  ggtitle("Saduria")

ggplot(bind_rows(p, d), aes(year, est, color = source)) +
  geom_point(size = 4) + 
  geom_errorbar(aes(ymin = est-sd, ymax = est+sd)) + 
  geom_line(size = 1) + 
  facet_wrap(~ SubDiv, scales = "free", ncol = 4) +
  theme_classic(base_size = 15) + 
  ggtitle("Saduria")

ggplot(bind_rows(p, d), aes(year, est, color = source)) +
  geom_point(size = 4) + 
  geom_line(size = 1) + 
  facet_wrap(~ SubDiv, scales = "free", ncol = 4) +
  theme_classic(base_size = 15) + 
  ggtitle("Saduria")
```

# Polychaeta
Read data

```{r pol data, message=FALSE}
# pol <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/bentfish/main/data/for_analysis/polychaeta.csv") %>%
#   dplyr::select(-X1)
pol <- read.csv("data/for_analysis/polychaeta.csv") %>% dplyr::select(-X)

pol <- pol %>%
  filter(year > 1985) %>% 
  mutate(SubDiv = as.factor(SubDiv)) %>% 
  filter(SubDiv %in% c(25, 27, 28)) %>% 
  mutate(year_f = as.factor(year_f),
         depth_scaled_sq = depth_scaled*depth_scaled)

# Split data by subdivison
pol25 <- pol %>% filter(SubDiv == "25")
pol25 %>%
  group_by(year) %>%
  summarize(n = n()) %>%
  ggplot(., aes(year, n)) + geom_bar(stat = "identity")

ggplot(pol25, aes(lon, lat, color = biomass)) + geom_point()
ggplot(pol25, aes(biomass)) + geom_histogram()

pol27 <- pol %>% filter(SubDiv == "27" & lon < 17.4)
pol27 %>%
  group_by(year) %>%
  summarize(n = n()) %>%
  ggplot(., aes(year, n)) + geom_bar(stat = "identity")

ggplot(pol27, aes(lon, lat, color = biomass)) + geom_point()
ggplot(pol27, aes(biomass)) + geom_histogram()

pol28 <- pol %>% filter(SubDiv == "28")
pol28 %>%
  group_by(year) %>%
  summarize(n = n()) %>%
  ggplot(., aes(year, n)) + geom_bar(stat = "identity")

ggplot(pol28, aes(biomass)) + geom_histogram()

pol28 <- pol28 %>% filter(year > 2005)

# remove data points with more than 80 g/m2
pol25 <- pol25 %>% filter(biomass < 100)
pol27 <- pol27 %>% filter(biomass < 100)
pol28 <- pol28 %>% filter(biomass < 80)
```

Read and crop pred grid to match polychaeta

Fit models using `sdmTMB` assuming a Tweedie distribution and biomass as the response.

Make spde mesh

```{r make pol barrier spde mesh, results='hide', cache=TRUE, message=FALSE}
# Non-island version
pol_spde25 <- make_mesh(data = pol25, xy_cols = c("lon", "lat"), n_knots = 50, type = "kmeans", seed = 42)

pol_spde27 <- make_mesh(data = pol27, xy_cols = c("lon", "lat"), n_knots = 45, type = "kmeans", seed = 42)

pol_spde28 <- make_mesh(data = pol28, xy_cols = c("lon", "lat"), n_knots = 20, type = "kmeans", seed = 42)
```

Fit the models    

```{r fit pol tweedie model, cache=TRUE}
m_pol25 <- sdmTMB(formula = biomass ~ year_f - 1, data = pol25,
                  time = "year", spde = pol_spde25, family = tweedie(link = "log"),
                  ar1_fields = FALSE, include_spatial = TRUE, spatial_trend = FALSE, spatial_only = TRUE)

m_pol27 <- sdmTMB(formula = biomass ~ year_f - 1, data = pol27,
                  time = "year", spde = pol_spde27, family = tweedie(link = "log"),
                  ar1_fields = FALSE, include_spatial = TRUE, spatial_trend = FALSE, spatial_only = TRUE)

m_pol28 <- sdmTMB(formula = biomass ~ year_f - 1, data = pol28,
                  time = "year", spde = pol_spde28, family = tweedie(link = "log"),
                  ar1_fields = FALSE, include_spatial = TRUE, spatial_trend = FALSE, spatial_only = TRUE)

# Check residuals
pol25$residuals <- residuals(m_pol25)
pol27$residuals <- residuals(m_pol27)
pol28$residuals <- residuals(m_pol28)

# Pretty good!
qqnorm(pol25$residuals); abline(a = 0, b = 1)
qqnorm(pol27$residuals); abline(a = 0, b = 1)
qqnorm(pol28$residuals); abline(a = 0, b = 1)

# Plot the marginal effect of depth:
# SD 25
# nd_pol25 <- data.frame(depth_scaled = seq(min(pol25$depth_scaled), max(pol25$depth_scaled), length.out = 100))
# nd_pol25$depth_scaled_sq <- nd_pol25$depth_scaled*nd_pol25$depth_scaled
# nd_pol25$year <- as.integer(max(pol25$year))
# nd_pol25$year_f <- factor(nd_pol25$year)
# 
# p_pol25 <- predict(m_pol25, newdata = nd_pol25, se_fit = TRUE, re_form = NA)
# 
# ggplot(p_pol25, aes(depth_scaled, exp(est),
#   ymin = exp(est - 1.96 * est_se), ymax = exp(est + 1.96 * est_se))) +
#   geom_line() + geom_ribbon(alpha = 0.4)
# 
# # SD 27
# nd_pol27 <- data.frame(depth_scaled = seq(min(pol27$depth_scaled), max(pol27$depth_scaled), length.out = 100))
# nd_pol27$depth_scaled_sq <- nd_pol27$depth_scaled*nd_pol27$depth_scaled
# nd_pol27$year <- as.integer(max(pol27$year))
# nd_pol27$year_f <- factor(nd_pol27$year)
# 
# p_pol27 <- predict(m_pol27, newdata = nd_pol27, se_fit = TRUE, re_form = NA)
# 
# ggplot(p_pol27, aes(depth_scaled, exp(est),
#   ymin = exp(est - 1.96 * est_se), ymax = exp(est + 1.96 * est_se))) +
#   geom_line() + geom_ribbon(alpha = 0.4)
# 
# # SD 28
# nd_pol28 <- data.frame(depth_scaled = seq(min(pol28$depth_scaled), max(pol28$depth_scaled), length.out = 100))
# nd_pol28$depth_scaled_sq <- nd_pol28$depth_scaled*nd_pol28$depth_scaled
# nd_pol28$year <- as.integer(max(pol28$year))
# nd_pol28$year_f <- factor(nd_pol28$year)
# 
# p_pol28 <- predict(m_pol28, newdata = nd_pol28, se_fit = TRUE, re_form = NA)
# 
# ggplot(p_pol28, aes(depth_scaled, exp(est),
#   ymin = exp(est - 1.96 * est_se), ymax = exp(est + 1.96 * est_se))) +
#   geom_line() + geom_ribbon(alpha = 0.4)
```

Read in the prediction grids
Read and crop pred grid to match polycheate

```{r crop pred grid for polycheate, message=FALSE, warning=FALSE}
## Read the prediction grid
# pred_grid <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/bentfish/main/data/for_analysis/pred_grid.csv")
pred_grid <- read.csv("data/for_analysis/pred_grid.csv")

pred_grid <- pred_grid %>%
  mutate(depth_scaled = (depth - mean(pol$depth))/sd(pol$depth)) %>% 
  mutate(depth_scaled_sq = depth_scaled*depth_scaled) %>% 
  mutate(year_f = factor(year),
         year = as.integer(year))

# Subset the prediction grid to match polycheate data
pred_grid_pol25 <- pred_grid %>% filter(SubDiv == "25" & year %in% c(unique(pol25$year)))
pred_grid_pol27 <- pred_grid %>% filter(SubDiv == "27" & year %in% c(unique(pol27$year)))
pred_grid_pol28 <- pred_grid %>% filter(SubDiv == "28" & year %in% c(unique(pol28$year)))

tf_pol25 <- exclude.too.far(pred_grid_pol25$lon, pred_grid_pol25$lat, pol$lon, pol$lat, 0.05) # 0.03 seems reasonable
tf_pol27 <- exclude.too.far(pred_grid_pol27$lon, pred_grid_pol27$lat, pol$lon, pol$lat, 0.05) # 0.03 seems reasonable
tf_pol28 <- exclude.too.far(pred_grid_pol28$lon, pred_grid_pol28$lat, pol$lon, pol$lat, 0.05) # 0.03 seems reasonable

# Filter the grid points that are not too far from the data
pred_grid_pol25$too_far <- tf_pol25
pred_grid_pol27$too_far <- tf_pol27
pred_grid_pol28$too_far <- tf_pol28

# Plot again
pred_grid_pol25 %>%
  filter(too_far == FALSE) %>%
  ggplot(., aes(lon, lat, fill = factor(SubDiv))) +
  geom_raster() +
  geom_point(data = pol25, aes(lon, lat), color = "black", size = 0.5) +
  NULL

pred_grid_pol27 %>%
  filter(too_far == FALSE) %>%
  ggplot(., aes(lon, lat, fill = factor(SubDiv))) +
  geom_raster() +
  geom_point(data = pol27, aes(lon, lat), color = "black", size = 0.5) +
  NULL

pred_grid_pol28 %>%
  filter(too_far == FALSE) %>%
  ggplot(., aes(lon, lat, fill = factor(SubDiv))) +
  geom_raster() +
  geom_point(data = pol28, aes(lon, lat), color = "black", size = 0.5) +
  NULL
```

Now calculate the index by year and sub area

```{r calculate pol index, cache=TRUE, message=FALSE}
# From these models, predict annual biomass for each sub area
# sort(unique(pol$SubDiv))

# Sub division 25
pol_preds25 <- predict(m_pol25, pred_grid_pol25, return_tmb_object = TRUE) # Predict using the grid for subarea

pol_ind25 <- get_index(pol_preds25, bias_correct = T) # Get the index (sum of all cells in the pred grid)

# The index is now the sum of grid cells in g/area of a grid cell
# First multiply by the area of a grid cell to get raw g
ncells25 <- pred_grid_pol25 %>% filter(year == 2015 & SubDiv == 25) %>% nrow()
# Then divide that by the total area

# This is the same as simply dividing by ncells25
pol_ind25 <- pol_ind25 %>% mutate(est = est/ncells25, lwr = lwr/ncells25, upr = upr/ncells25, SubDiv = 25)


# Sub division 27
pol_preds27 <- predict(m_pol27, pred_grid_pol27, return_tmb_object = TRUE) # Predict using the grid for subarea

pol_ind27 <- get_index(pol_preds27, bias_correct = F) # Get the index (sum of all cells in the pred grid)

# The index is now the sum of grid cells in g/area of a grid cell
# First multiply by the area of a grid cell to get raw g
ncells27 <- pred_grid_pol27 %>% filter(year == 2015 & SubDiv == 27) %>% nrow()
# Then divide that by the total area

# This is the same as simply dividing by ncells27
pol_ind27 <- pol_ind27 %>% mutate(est = est/ncells27, lwr = lwr/ncells27, upr = upr/ncells27, SubDiv = 27)


# Sub division 28
pol_preds28 <- predict(m_pol28, pred_grid_pol28, return_tmb_object = TRUE) # Predict using the grid for subarea

pol_ind28 <- get_index(pol_preds28, bias_correct = T) # Get the index (sum of all cells in the pred grid)

# The index is now the sum of grid cells in g/area of a grid cell
# First multiply by the area of a grid cell to get raw g
ncells28 <- pred_grid_pol28 %>% filter(year == 2015 & SubDiv == 28) %>% nrow()
# Then divide that by the total area

# This is the same as simply dividing by ncells28
pol_ind28 <- pol_ind28 %>% mutate(est = est/ncells28, lwr = lwr/ncells28, upr = upr/ncells28, SubDiv = 28)


# Merge prediction-data
pol_preds <- bind_rows(pol_ind25, pol_ind27, pol_ind28) %>% mutate(SubDiv = factor(SubDiv))

# Compare to data quickly
p <- pol_preds %>%
  dplyr::select(year, est, upr, lwr, SubDiv) %>%
  mutate(source = "pred") %>% 
  arrange(year, SubDiv)

d <- pol %>%
  group_by(year, SubDiv) %>% 
  summarise(est = mean(biomass),
            sd = sd(biomass)) %>%
  mutate(source = "data") %>% 
  arrange(year, SubDiv)

ggplot(bind_rows(p, d), aes(year, est, color = source)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = est-sd, ymax = est+sd)) + 
  geom_line(size = 1) + 
  facet_wrap(~ SubDiv, scales = "free", ncol = 4) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.4) + 
  theme_classic(base_size = 15) + 
  ggtitle("Polychaeta")

ggplot(bind_rows(p, d), aes(year, est, color = source)) +
  geom_point(size = 4) + 
  geom_errorbar(aes(ymin = est-sd, ymax = est+sd)) + 
  geom_line(size = 1) + 
  facet_wrap(~ SubDiv, scales = "free", ncol = 4) +
  theme_classic(base_size = 15) + 
  ggtitle("Polychaeta")

ggplot(bind_rows(p, d), aes(year, est, color = source)) +
  geom_point(size = 4) + 
  geom_line(size = 1) + 
  facet_wrap(~ SubDiv, scales = "free", ncol = 4) +
  theme_classic(base_size = 15) + 
  ggtitle("Polychaeta")
```

# Limecola
Read data

```{r lim data, message=FALSE}
# lim <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/bentfish/main/data/for_analysis/limecola.csv") %>%
#   dplyr::select(-X1)
lim <- read.csv("data/for_analysis/limecola.csv") %>% dplyr::select(-X)

lim <- lim %>%
  filter(year > 1985) %>% 
  mutate(SubDiv = as.factor(SubDiv)) %>% 
  filter(SubDiv %in% c(25, 27, 28)) %>% 
  mutate(year_f = as.factor(year_f),
         depth_scaled_sq = depth_scaled*depth_scaled)

# Split data by subdivison
lim25 <- lim %>% filter(SubDiv == "25")
lim25 %>%
  group_by(year) %>%
  summarize(n = n()) %>%
  ggplot(., aes(year, n)) + geom_bar(stat = "identity")

ggplot(lim25, aes(lon, lat, color = biomass)) + geom_point()
ggplot(lim25, aes(biomass)) + geom_histogram()

lim27 <- lim %>% filter(SubDiv == "27" & lon < 17.4)
lim27 %>%
  group_by(year) %>%
  summarize(n = n()) %>%
  ggplot(., aes(year, n)) + geom_bar(stat = "identity")

ggplot(lim27, aes(lon, lat, color = biomass)) + geom_point()
ggplot(lim27, aes(biomass)) + geom_histogram()

lim28 <- lim %>% filter(SubDiv == "28")
lim28 %>%
  group_by(year) %>%
  summarize(n = n()) %>%
  ggplot(., aes(year, n)) + geom_bar(stat = "identity")

ggplot(lim28, aes(biomass)) + geom_histogram()

lim28 <- lim28 %>% filter(year > 2005)

# remove data points with more than 80 g/m2
lim25 <- lim25 %>% filter(biomass < 300)
lim27 <- lim27 %>% filter(biomass < 300)
lim28 <- lim28 %>% filter(biomass < 190)
```

Read and crop pred grid to match limecola

Fit models using `sdmTMB` assuming a Tweedie distribution and biomass as the response.

Make spde mesh

```{r make lim barrier spde mesh, results='hide', cache=TRUE, message=FALSE}
# Non-island version
lim_spde25 <- make_mesh(data = lim25, xy_cols = c("lon", "lat"), n_knots = 80, type = "kmeans", seed = 42)

lim_spde27 <- make_mesh(data = lim27, xy_cols = c("lon", "lat"), n_knots = 50, type = "kmeans", seed = 42)

lim_spde28 <- make_mesh(data = lim28, xy_cols = c("lon", "lat"), n_knots = 40, type = "kmeans", seed = 42)
```

Fit the models

```{r fit lim tweedie model, cache=TRUE}
m_lim25 <- sdmTMB(formula = biomass ~ year_f - 1, data = lim25,
                  time = "year", spde = lim_spde25, family = tweedie(link = "log"),
                  ar1_fields = FALSE, include_spatial = TRUE, spatial_trend = FALSE, spatial_only = TRUE)

m_lim27 <- sdmTMB(formula = biomass ~ year_f - 1, data = lim27,
                  time = "year", spde = lim_spde27, family = tweedie(link = "log"),
                  ar1_fields = FALSE, include_spatial = TRUE, spatial_trend = FALSE, spatial_only = TRUE)

m_lim28 <- sdmTMB(formula = biomass ~ year_f - 1, data = lim28,
                  time = "year", spde = lim_spde28, family = tweedie(link = "log"),
                  ar1_fields = FALSE, include_spatial = TRUE, spatial_trend = FALSE, spatial_only = TRUE)

# Check residuals
lim25$residuals <- residuals(m_lim25)
lim27$residuals <- residuals(m_lim27)
lim28$residuals <- residuals(m_lim28)

# Pretty good!
qqnorm(lim25$residuals); abline(a = 0, b = 1)
qqnorm(lim27$residuals); abline(a = 0, b = 1)
qqnorm(lim28$residuals); abline(a = 0, b = 1)

# Plot the marginal effect of depth:
# SD 25
# nd_lim25 <- data.frame(depth_scaled = seq(min(lim25$depth_scaled), max(lim25$depth_scaled), length.out = 100))
# nd_lim25$depth_scaled_sq <- nd_lim25$depth_scaled*nd_lim25$depth_scaled
# nd_lim25$year <- as.integer(max(lim25$year))
# nd_lim25$year_f <- factor(nd_lim25$year)
# 
# p_lim25 <- predict(m_lim25, newdata = nd_lim25, se_fit = TRUE, re_form = NA)
# 
# ggplot(p_lim25, aes(depth_scaled, exp(est),
#   ymin = exp(est - 1.96 * est_se), ymax = exp(est + 1.96 * est_se))) +
#   geom_line() + geom_ribbon(alpha = 0.4)
# 
# # SD 27
# nd_lim27 <- data.frame(depth_scaled = seq(min(lim27$depth_scaled), max(lim27$depth_scaled), length.out = 100))
# nd_lim27$depth_scaled_sq <- nd_lim27$depth_scaled*nd_lim27$depth_scaled
# nd_lim27$year <- as.integer(max(lim27$year))
# nd_lim27$year_f <- factor(nd_lim27$year)
# 
# p_lim27 <- predict(m_lim27, newdata = nd_lim27, se_fit = TRUE, re_form = NA)
# 
# ggplot(p_lim27, aes(depth_scaled, exp(est),
#   ymin = exp(est - 1.96 * est_se), ymax = exp(est + 1.96 * est_se))) +
#   geom_line() + geom_ribbon(alpha = 0.4)
# 
# # SD 28
# nd_lim28 <- data.frame(depth_scaled = seq(min(lim28$depth_scaled), max(lim28$depth_scaled), length.out = 100))
# nd_lim28$depth_scaled_sq <- nd_lim28$depth_scaled*nd_lim28$depth_scaled
# nd_lim28$year <- as.integer(max(lim28$year))
# nd_lim28$year_f <- factor(nd_lim28$year)
# 
# p_lim28 <- predict(m_lim28, newdata = nd_lim28, se_fit = TRUE, re_form = NA)
# 
# ggplot(p_lim28, aes(depth_scaled, exp(est),
#   ymin = exp(est - 1.96 * est_se), ymax = exp(est + 1.96 * est_se))) +
#   geom_line() + geom_ribbon(alpha = 0.4)
```

Read in the prediction grids
Read and crop pred grid to match limecola

```{r crop pred grid for limecola, message=FALSE, warning=FALSE}
## Read the prediction grid
# pred_grid <- readr::read_csv("https://raw.githubusercontent.com/maxlindmark/bentfish/main/data/for_analysis/pred_grid.csv")
pred_grid <- read.csv("data/for_analysis/pred_grid.csv")

pred_grid <- pred_grid %>%
  mutate(depth_scaled = (depth - mean(lim$depth))/sd(lim$depth)) %>% 
  mutate(depth_scaled_sq = depth_scaled*depth_scaled) %>% 
  mutate(year_f = factor(year),
         year = as.integer(year))

# Subset the prediction grid to match limecola data
pred_grid_lim25 <- pred_grid %>% filter(SubDiv == "25" & year %in% c(unique(lim25$year)))
pred_grid_lim27 <- pred_grid %>% filter(SubDiv == "27" & year %in% c(unique(lim27$year)))
pred_grid_lim28 <- pred_grid %>% filter(SubDiv == "28" & year %in% c(unique(lim28$year)))

tf_lim25 <- exclude.too.far(pred_grid_lim25$lon, pred_grid_lim25$lat, lim$lon, lim$lat, 0.05) # 0.03 seems reasonable
tf_lim27 <- exclude.too.far(pred_grid_lim27$lon, pred_grid_lim27$lat, lim$lon, lim$lat, 0.05) # 0.03 seems reasonable
tf_lim28 <- exclude.too.far(pred_grid_lim28$lon, pred_grid_lim28$lat, lim$lon, lim$lat, 0.05) # 0.03 seems reasonable

# Filter the grid points that are not too far from the data
pred_grid_lim25$too_far <- tf_lim25
pred_grid_lim27$too_far <- tf_lim27
pred_grid_lim28$too_far <- tf_lim28

# Plot again
pred_grid_lim25 %>%
  filter(too_far == FALSE) %>%
  ggplot(., aes(lon, lat, fill = factor(SubDiv))) +
  geom_raster() +
  geom_point(data = lim25, aes(lon, lat), color = "black", size = 0.5) +
  NULL

pred_grid_lim27 %>%
  filter(too_far == FALSE) %>%
  ggplot(., aes(lon, lat, fill = factor(SubDiv))) +
  geom_raster() +
  geom_point(data = lim27, aes(lon, lat), color = "black", size = 0.5) +
  NULL

pred_grid_lim28 %>%
  filter(too_far == FALSE) %>%
  ggplot(., aes(lon, lat, fill = factor(SubDiv))) +
  geom_raster() +
  geom_point(data = lim28, aes(lon, lat), color = "black", size = 0.5) +
  NULL
```

Now calculate the index by year and sub area

```{r calculate lim index, cache=TRUE, message=FALSE}
# From these models, predict annual biomass for each sub area
# sort(unique(lim$SubDiv))

# Sub division 25
lim_preds25 <- predict(m_lim25, pred_grid_lim25, return_tmb_object = TRUE) # Predict using the grid for subarea

lim_ind25 <- get_index(lim_preds25, bias_correct = T) # Get the index (sum of all cells in the pred grid)

# The index is now the sum of grid cells in g/area of a grid cell
# First multiply by the area of a grid cell to get raw g
ncells25 <- pred_grid_lim25 %>% filter(year == 2015 & SubDiv == 25) %>% nrow()
# Then divide that by the total area

# This is the same as simply dividing by ncells25
lim_ind25 <- lim_ind25 %>% mutate(est = est/ncells25, lwr = lwr/ncells25, upr = upr/ncells25, SubDiv = 25)


# Sub division 27
lim_preds27 <- predict(m_lim27, pred_grid_lim27, return_tmb_object = TRUE) # Predict using the grid for subarea

lim_ind27 <- get_index(lim_preds27, bias_correct = F) # Get the index (sum of all cells in the pred grid)

# The index is now the sum of grid cells in g/area of a grid cell
# First multiply by the area of a grid cell to get raw g
ncells27 <- pred_grid_lim27 %>% filter(year == 2015 & SubDiv == 27) %>% nrow()
# Then divide that by the total area

# This is the same as simply dividing by ncells27
lim_ind27 <- lim_ind27 %>% mutate(est = est/ncells27, lwr = lwr/ncells27, upr = upr/ncells27, SubDiv = 27)


# Sub division 28
lim_preds28 <- predict(m_lim28, pred_grid_lim28, return_tmb_object = TRUE) # Predict using the grid for subarea

lim_ind28 <- get_index(lim_preds28, bias_correct = T) # Get the index (sum of all cells in the pred grid)

# The index is now the sum of grid cells in g/area of a grid cell
# First multiply by the area of a grid cell to get raw g
ncells28 <- pred_grid_lim28 %>% filter(year == 2015 & SubDiv == 28) %>% nrow()
# Then divide that by the total area

# This is the same as simply dividing by ncells28
lim_ind28 <- lim_ind28 %>% mutate(est = est/ncells28, lwr = lwr/ncells28, upr = upr/ncells28, SubDiv = 28)


# Merge prediction-data
lim_preds <- bind_rows(lim_ind25, lim_ind27, lim_ind28) %>% mutate(SubDiv = factor(SubDiv))

# Compare to data quickly
p <- lim_preds %>%
  dplyr::select(year, est, upr, lwr, SubDiv) %>%
  mutate(source = "pred") %>% 
  arrange(year, SubDiv)

d <- lim %>%
  group_by(year, SubDiv) %>% 
  summarise(est = mean(biomass),
            sd = sd(biomass)) %>%
  mutate(source = "data") %>% 
  arrange(year, SubDiv)
  
ggplot(bind_rows(p, d), aes(year, est, color = source)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = est-sd, ymax = est+sd)) + 
  geom_line(size = 1) + 
  facet_wrap(~ SubDiv, scales = "free", ncol = 4) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.4) + 
  theme_classic(base_size = 15) + 
  ggtitle("Limecola")

ggplot(bind_rows(p, d), aes(year, est, color = source)) +
  geom_point(size = 4) + 
  geom_errorbar(aes(ymin = est-sd, ymax = est+sd)) + 
  geom_line(size = 1) + 
  facet_wrap(~ SubDiv, scales = "free", ncol = 4) +
  theme_classic(base_size = 15) + 
  ggtitle("Limecola")

ggplot(bind_rows(p, d), aes(year, est, color = source)) +
  geom_point(size = 4) + 
  geom_line(size = 1) + 
  facet_wrap(~ SubDiv, scales = "free", ncol = 4) +
  theme_classic(base_size = 15) + 
  ggtitle("Limecola")
```

# Merge data to a full index
```{r merge all predictions}
sad_preds <- sad_preds %>%
  dplyr::select(year, SubDiv, est, lwr, upr) %>%
  mutate(species_group = "Saduria entomon", source = "model")

lim_preds <- lim_preds %>%
  dplyr::select(year, SubDiv, est, lwr, upr) %>%
  mutate(species_group = "Limecola balthica", source = "model")

pol_preds <- pol_preds %>%
  dplyr::select(year, SubDiv, est, lwr, upr) %>%
  mutate(species_group = "Polychaeta", source = "model")

pred_dat <- bind_rows(sad_preds, lim_preds, pol_preds)

# Now add in the data
dat <- bind_rows(sad25, sad27, sad28,
                 lim25, lim27, lim28,
                 pol25, pol27, pol28) %>% dplyr::select(year, SubDiv, species_group, biomass)

avg_dat <- dat %>%
  group_by(year, SubDiv, species_group) %>%
  summarize(est = mean(biomass)) %>%
  mutate(source = "data")

full_dat <- bind_rows(avg_dat, pred_dat)

# Plot
ggplot(full_dat, aes(year, est, color = factor(SubDiv), shape = source, linetype = source)) +
  geom_point(size = 3) +
  geom_line(size = 0.5) +
  scale_color_brewer(palette = "Dark2", name = "sub-area") +
  facet_wrap(~ species_group, scales = "free", ncol = 3) +
  theme_classic(base_size = 15) +
  theme(legend.position = "bottom")
ggsave("figures/long_index_data.png", width = 9, height = 9, dpi = 600)

# Now without data
full_dat %>%
  filter(source == "model") %>%
  ggplot(., aes(year, est, color = factor(SubDiv), fill = factor(SubDiv))) +
  geom_point(size = 4) +
  geom_line(size = 1) +
  scale_color_brewer(palette = "Dark2", name = "sub-area") +
  scale_fill_brewer(palette = "Dark2", name = "sub-area") +
  facet_wrap(~ species_group, scales = "free", ncol = 3) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
  theme_classic(base_size = 15) +
  theme(legend.position = "bottom")

full_dat %>%
  filter(source == "model") %>%
  ggplot(., aes(year, est, color = factor(SubDiv), fill = factor(SubDiv))) +
  geom_point(size = 4) +
  geom_line(size = 1) +
  scale_color_brewer(palette = "Dark2", name = "sub-area") +
  scale_fill_brewer(palette = "Dark2", name = "sub-area") +
  facet_wrap(~ species_group, scales = "free", ncol = 3) +
  theme_classic(base_size = 15) +
  theme(legend.position = "bottom")

ggsave("figures/long_index_ci.png", width = 9, height = 9, dpi = 600)

write.csv(full_dat, "output/long_benthic_indicies.csv")
```
