---
title: "Joining benthic data"
author: "Max Lindmark"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 12,
  fig.asp = 0.618,
  fig.align ='center'
)
```

## Overview

The goal with this script is to merge all benthic datasets to a complete
one in a tidy format, where 1 row = 1 sample (by species).

#### Benthic samples

-   What happens here is that I first do a full join for abundance and
    biomass data. Next I do a left join to add in taxonomic information
    to easily navigate between the species and groups present in the
    abundance and biomass data. Lastly, I do a left join with the
    provinfo data to provide each sample ID with geographic coordinates.

After that I show some basic exploratory plots that can be used as
templates for exploring e.g. certain species or years in more detail,
simply by applying different filters and data manipulations.

The data were sent out by Mattias Sköld, and are also stored here in
this RProject, as they came.

Please help me out finding all the errors and loopholes! I have not
worked with this data before.

That said, let's start. First, read in some helpful packages:

```{r read libs, warning=FALSE, message=FALSE}
rm(list = ls()) # Clear console history (restart r for complete fresh start)

# Load libraries (install first if needed)
library(tidyverse)
library(tidylog)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(forcats)
library(gapminder)
library(viridis)
library(raster)
library(ncdf4)
library(chron)
library(mapplots)

theme_set(theme_classic())

# Quick and dirty way to plot spatial data
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

xlim <- c(12.99, 19.97) # c(min(dat$Longitud), max(dat$Longitud))
ylim <- c(55.12, 60.39) # c(min(dat$Latitud), max(dat$Latitud))

# Define a function for plotting maps...
theme_map <- function () { 
  theme(
    text = element_text(size = 8),
    legend.position = "bottom", 
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8),
    axis.text = element_text(size = 6))
}
```

Next start by collating and cleaning the benthic data

## Benthic data

Next we read in the data (abundans, biomass and provinfo) as they came
and do so minor tidying up:

```{r read data, warning=FALSE, message=FALSE}
# I will read them as they came - tab delimited .txt files. 
abun <- read_delim("data/benthic_data/abundans.txt",
                    skip = 1, delim = "\t") # skip 1st row

biom <- read_delim("data/benthic_data/biomassa.txt",
                    skip = 1, delim = "\t") # skip 1st row

# Remove the 4th column because it's empty in the data and only added with blank cells
abun <- abun %>% dplyr::select(-"X4")
biom <- biom %>% dplyr::select(-"X4")

# The prov_info file does not have a header row
# Also, open it first in Firefox, click view and find it's encoding. On my machine I
# need to specify that manually. Therefore I use the function read.delim not read_delim
# and convert it to a tibble afterwards
prov_info <- read.delim("data/benthic_data/provinfo.txt",
                        sep = "\t", fileEncoding= "windows-1252") 

prov <- tibble(prov_info)

# Inspect the structure of the datasets 
str(abun)
str(biom)

# Biomass is continuous and separated by comma as a character. as.numeric won't do the trick
head(biom$biomassa)
biom$biomassa <- as.numeric(sub(",", ".", biom$biomassa, fixed = TRUE))
head(biom$biomassa)

str(prov)

# Convert date in prov to date type
prov$Datum <- as.Date(prov$Datum)
```

At this point, the data are already 1 row = 1 sample. It might come from
the same station, however, so one idea is to aggregate this. If that's
appropriate or not depends on how we use the data. I will not do it here
for now.

Now let's join the different datasets.

Start with biomass and abundance and join by a unique id in order to get
a single data.

```{r join data, warning=FALSE, message=FALSE}
# Which columns are *not* in the other data set?
# The only columns that differ are "antal" and "biomass", so let's just do a left_join
colnames(abun)[!colnames(abun) %in% colnames(biom)]
colnames(biom)[!colnames(biom) %in% colnames(abun)]

# Ensure that ID does not restart over year etc.
length(unique(abun$provID))
length(unique(paste(abun$provID, abun$artal)))

length(unique(biom$provID))
length(unique(paste(biom$provID, biom$artal)))

# The data are of different lengths. There are more IDs in the abundance data

# Create a new ID column that is the provID + taxa so that I can match in the data
abun$id_yeartax <- paste(abun$provID, abun$taxa, sep = ".")
biom$id_yeartax <- paste(biom$provID, biom$taxa, sep = ".")

# Check that we only have one row per id!
biom %>% 
  group_by(id_yeartax) %>% 
  summarise(n = n()) %>% 
  distinct(n)

abun %>% 
  group_by(id_yeartax) %>% 
  summarise(n = n()) %>% 
  distinct(n)

# Yes!

# Now we can do the full_join to add biomass data as a column to the abundance dataset.
# We don't do a left_join here because there are some rows that do not exist in the abundance data and vice versa (check below which ID's are *not* in the other data set).
id_not_in_biom <- abun$id_yeartax[!abun$id_yeartax %in% biom$id_yeartax]
id_not_in_abun <- biom$id_yeartax[!biom$id_yeartax %in% abun$id_yeartax]

length(id_not_in_biom)
length(id_not_in_abun)

# The new data will be called "dat". First drop some non-needed columns in the biomass data
biom_small <- biom %>% dplyr::select(biomassa, id_yeartax)

dat <- full_join(abun, biom_small, by = "id_yeartax")
# Now we got 98,044 rows, of which 95,925 are common id's (98,044 - 95,925 = length(id_not_in_biom) + length(id_not_in_abun))

# Check it worked...
nrow(dat)
colnames(dat)

# Check the 5th ID that is in the abun but not in the biom
test_id <- id_not_in_biom[5]

# Check in the abun data
abun %>% filter(id_yeartax == test_id)
biom %>% filter(id_yeartax == test_id)
dat %>% filter(id_yeartax == test_id)

# Looks correct!
```

Now I want the taxonomic information from the file "Mattias BenthFish
artlista.xlsx". The goal is to have the columns "taxonID", "Stam",
"Klass", "Ordning", "Familj" in the abun and biom data so that rookies
with no taxonomic skills whatsoever like myself more easily can get an
overview.

```{r add taxa, warning=FALSE, message=FALSE}
taxon_info <- readxl::read_xlsx("data/benthic_data/Mattias BenthFish artlista.xlsx")
taxon_info

# Now do a left_join to add the full taxon information to dat from taxon_info
# (First copy the taxa column so that I can verify it after)
dat$taxa_original <- dat$taxa
dat <- left_join(dat, taxon_info, by = "taxa")

# Check it worked
dat %>% 
  dplyr::select(c(taxa_original, colnames(taxon_info))) %>% 
  dplyr::select(-c(Synonymer, Auktor, taxonID))  # Remove these for checking only
```

Lastly, we want provinfo data to be joined in as well to get e.g.
coordinates

```{r add sample info, warning=FALSE, message=FALSE}
colnames(prov)
colnames(dat)

# We will save all information from here as I don't know what all columns mean. Here we use the provID as the key
prov <- prov %>% dplyr::select(-artal, -djup, -Datum) # To not get double columns in new data
dat <- left_join(dat, prov, by = "provID")

str(dat)

# Need to change the classes of coordinates
head(dat$Latitud)
dat$Latitud <- as.numeric(sub(",", ".", dat$Latitud, fixed = TRUE))
head(dat$Latitud)

dat$Longitud <- as.numeric(sub(",", ".", dat$Longitud, fixed = TRUE))

# Remove NA-coordinates (22... hmm, could be ones that were not in the abundance data but not in the biomass data. Doesn't matter though!)
dat <- dat %>% drop_na(c(Longitud, Latitud))
```

Split the date column so that it will be easier to by month subset later

```{r split date column, message = FALSE}
dat <- dat %>%
  separate(datum, c("Year", "Month", "Day"), sep = "-") %>%
  mutate(Year = as.numeric(Year),
         Month = as.numeric(Month),
         Day = as.numeric(Day))
```

Add in oceanographic variables. Start with oxygen

```{r oxygen, message=FALSE, results="hide"}
# # Oxygen 
# # Loop through each year and extract the oxygen levels
# # Downloaded from here: https://resources.marine.copernicus.eu/?option=com_csw&view=details&product_id=BALTICSEA_REANALYSIS_BIO_003_012
# # Extract raster points: https://gisday.wordpress.com/2014/03/24/extract-raster-values-from-points-using-r/comment-page-1/
# # https://rpubs.com/boyerag/297592
# # https://pjbartlein.github.io/REarthSysSci/netCDF.html#get-a-variable
# # Open the netCDF file
# ncin <- nc_open("data/NEMO_Nordic_SCOBI/dataset-reanalysis-scobi-monthlymeans_1610091357600.nc")
# 
# print(ncin)
# 
# # Get longitude and latitude
# lon <- ncvar_get(ncin,"longitude")
# nlon <- dim(lon)
# head(lon)
# 
# lat <- ncvar_get(ncin,"latitude")
# nlat <- dim(lat)
# head(lat)
# 
# # Get time
# time <- ncvar_get(ncin,"time")
# time
# 
# tunits <- ncatt_get(ncin,"time","units")
# nt <- dim(time)
# nt
# tunits
# 
# # Get oxygen
# dname <- "o2b"
# 
# oxy_array <- ncvar_get(ncin,dname)
# dlname <- ncatt_get(ncin,dname,"long_name")
# dunits <- ncatt_get(ncin,dname,"units")
# fillvalue <- ncatt_get(ncin,dname,"_FillValue")
# dim(oxy_array)
# 
# # Get global attributes
# title <- ncatt_get(ncin,0,"title")
# institution <- ncatt_get(ncin,0,"institution")
# datasource <- ncatt_get(ncin,0,"source")
# references <- ncatt_get(ncin,0,"references")
# history <- ncatt_get(ncin,0,"history")
# Conventions <- ncatt_get(ncin,0,"Conventions")
# 
# # Convert time: split the time units string into fields
# tustr <- strsplit(tunits$value, " ")
# tdstr <- strsplit(unlist(tustr)[3], "-")
# tmonth <- as.integer(unlist(tdstr)[2])
# tday <- as.integer(unlist(tdstr)[3])
# tyear <- as.integer(unlist(tdstr)[1])
# 
# # Here I deviate from the guide a little bit. Save this info:
# dates <- chron(time, origin = c(tmonth, tday, tyear))
# 
# # Crop the date variable
# months <- as.numeric(substr(dates, 2, 3))
# years <- as.numeric(substr(dates, 8, 9))
# years <- ifelse(years > 90, 1900 + years, 2000 + years)
# 
# # Replace netCDF fill values with NA's
# oxy_array[oxy_array == fillvalue$value] <- NA
# 
# # We only use Months 4-6 (quarter 2) in this analysis, so now we want to loop through each time step,
# # and if it is a good month save it as a raster.
# # First get the index of months that correspond to Q4
# months
# 
# index_keep <- which(months > 3 & months < 7)
# 
# # Quarter 2 by keeping months in index_keep
# oxy_q2 <- oxy_array[, , index_keep]
# 
# months_keep <- months[index_keep]
# 
# years_keep <- years[index_keep]
# 
# # Now we have an array with only Q2 data...
# # We need to now calculate the average within a year.
# # Get a sequence that takes every third value between 1: number of months (length)
# loop_seq <- seq(1, dim(oxy_q2)[3], by = 3)
# 
# # Create objects that will hold data
# dlist <- list()
# oxy_4 <- c()
# oxy_5 <- c()
# oxy_6 <- c()
# oxy_ave <- c()
# 
# # Loop through the vector sequence with every third value, then take the average of
# # three consecutive months (i.e. q2)
# for(i in loop_seq) {
#   
#   oxy_4 <- oxy_q2[, , (i)]
#   oxy_5 <- oxy_q2[, , (i + 1)]
#   oxy_6 <- oxy_q2[, , (i + 2)]
#   
#   oxy_ave <- (oxy_4 + oxy_5 + oxy_6) / 3
#   
#   list_pos <- ((i/3) - (1/3)) + 1 # to get index 1:n(years)
#   
#   dlist[[list_pos]] <- oxy_ave
#   
# }
# 
# # Now name the lists with the year:
# # First rename some variables
# dat$year <- dat$artal
# dat$lat <- dat$Latitud
# dat$lon <- dat$Longitud
# 
# names(dlist) <- unique(years_keep)
# 
# # Now I need to make a loop where I extract the raster value for each year...
# 
# # Filter years in the benthic data frame to only have the years I have oxygen for
# d_sub_oxy <- dat %>% filter(year %in% names(dlist)) %>% droplevels()
# 
# # Now keep only the unique combinations of lat lon and year
# d_sub_oxy <- d_sub_oxy %>% mutate(id_oxy = paste(lat, lon, year)) %>% distinct(id_oxy, .keep_all = TRUE)
# 
# # Create data holding object
# data_list <- list()
# 
# # Create factor year for indexing the list in the loop
# d_sub_oxy$year_f <- as.factor(d_sub_oxy$year)
# 
# # Loop through each year and extract raster values for the pred grid data points
# for(i in unique(d_sub_oxy$year_f)) {
#   
#   # Subset a year
#   oxy_slice <- dlist[[i]]
#   
#   # Create raster for that year (i)
#   r <- raster(t(oxy_slice), xmn = min(lon), xmx = max(lon), ymn = min(lat), ymx = max(lat),
#               crs = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
#   
#   # Flip...
#   r <- flip(r, direction = 'y')
#   
#   plot(r, main = i)
#   
#   # Filter the same year (i) in the pred-grid data and select only coordinates
#   d_slice <- d_sub_oxy %>% filter(year_f == i) %>% dplyr::select(lon, lat)
#   
#   # Make into a SpatialPoints object
#   data_sp <- SpatialPoints(d_slice)
#   
#   # Extract raster value (oxygen)
#   rasValue <- raster::extract(r, data_sp)
#   
#   # Now we want to plot the results of the raster extractions by plotting the pred-grid
#   # data points over a raster and saving it for each year.
#   # Make the SpatialPoints object into a raster again (for pl)
#   df <- as.data.frame(data_sp)
#   
#   # Add in the raster value in the df holding the coordinates for the pred-grid data
#   d_slice$oxy <- rasValue
#   
#   # Add in which year
#   d_slice$year <- i
#   
#   # Create a index for the data last where we store all years (because our loop index
#   # i is not continuous, we can't use it directly)
#   index <- as.numeric(d_slice$year)[1] - 1992
#   
#   # Add each years' data in the list
#   data_list[[index]] <- d_slice
#   
# }
# 
# # Now create a data frame from the list of all annual values
# pred_grid_oxy <- dplyr::bind_rows(data_list)
# 
# lims <- pred_grid_oxy %>% drop_na(oxy) %>% summarise(min = min(oxy),
#                                                      max = max(oxy))
# 
# # Plot and compare with rasters
# ggplot(pred_grid_oxy, aes(lon, lat, color = oxy)) + 
#   facet_wrap(~year) +
#   scale_colour_gradientn(colours = rev(terrain.colors(10)),
#                          limits = c(lims$min, lims$max)) +
#   geom_sf(data = world, inherit.aes = F, size = 0.2, alpha = 0) +
#   coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
#   geom_point(size = 0.1) +
#   NULL
#   
# # Left join in with the benthic again
# # First make the same selection of years in pred-grid 2
# # sort(unique(dat$year))
# 
# pred_grid_oxy <- pred_grid_oxy %>% mutate(id_oxy = paste(year, lon, lat)) %>% dplyr::select(-year, -lon, -lat)
# 
# dat <- dat %>% mutate(id_oxy = paste(year, lon, lat))
# 
# # Add in oxygen
# nrow(dat)
# dat <- left_join(dat, pred_grid_oxy)
# nrow(dat)
#      
# ggplot(dat, aes(lon, lat, color = oxy)) + 
#   facet_wrap(~year) +
#   scale_colour_gradientn(colours = rev(terrain.colors(10)),
#                          limits = c(lims$min, lims$max)) +
#   geom_sf(data = world, inherit.aes = F, size = 0.2, alpha = 0) +
#   coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
#   geom_point(size = 0.1) +
#   NULL
```

Now do temperature

```{r temperature, message=FALSE, results="hide"}
# # Open the netCDF file
# ncin <- nc_open("data/NEMO_Nordic_SCOBI/dataset-reanalysis-nemo-monthlymeans_1608127623694.nc")
# 
# print(ncin)
# 
# # Get longitude and latitude
# lon <- ncvar_get(ncin,"longitude")
# nlon <- dim(lon)
# head(lon)
# 
# lat <- ncvar_get(ncin,"latitude")
# nlat <- dim(lat)
# head(lat)
# 
# # Get time
# time <- ncvar_get(ncin,"time")
# time
# 
# tunits <- ncatt_get(ncin,"time","units")
# nt <- dim(time)
# nt
# tunits
# 
# # Get temperature
# dname <- "bottomT"
# 
# temp_array <- ncvar_get(ncin,dname)
# dlname <- ncatt_get(ncin,dname,"long_name")
# dunits <- ncatt_get(ncin,dname,"units")
# fillvalue <- ncatt_get(ncin,dname,"_FillValue")
# dim(temp_array)
# 
# # Get global attributes
# title <- ncatt_get(ncin,0,"title")
# institution <- ncatt_get(ncin,0,"institution")
# datasource <- ncatt_get(ncin,0,"source")
# references <- ncatt_get(ncin,0,"references")
# history <- ncatt_get(ncin,0,"history")
# Conventions <- ncatt_get(ncin,0,"Conventions")
# 
# # Convert time: split the time units string into fields
# tustr <- strsplit(tunits$value, " ")
# tdstr <- strsplit(unlist(tustr)[3], "-")
# tmonth <- as.integer(unlist(tdstr)[2])
# tday <- as.integer(unlist(tdstr)[3])
# tyear <- as.integer(unlist(tdstr)[1])
# 
# # Here I deviate from the guide a little bit. Save this info:
# dates <- chron(time, origin = c(tmonth, tday, tyear))
# 
# # Crop the date variable
# months <- as.numeric(substr(dates, 2, 3))
# years <- as.numeric(substr(dates, 8, 9))
# years <- ifelse(years > 90, 1900 + years, 2000 + years)
# 
# # Replace netCDF fill values with NA's
# temp_array[temp_array == fillvalue$value] <- NA
# 
# # We only use Quarter 2 in this analysis, so now we want to loop through each time step,
# # and if it is a good month save it as a raster.
# # First get the index of months that correspond to Q2
# months
# 
# index_keep <- which(months > 3 & months < 7)
# 
# # Quarter 2 by keeping months in index_keep
# temp_q2 <- temp_array[, , index_keep]
# 
# months_keep <- months[index_keep]
# 
# years_keep <- years[index_keep]
# 
# # Now we have an array with only Q2 data...
# # We need to now calculate the average within a year.
# # Get a sequence that takes every third value between 1: number of months (length)
# loop_seq <- seq(1, dim(temp_q2)[3], by = 3)
# 
# # Create objects that will hold data
# dlist <- list()
# temp_4 <- c()
# temp_5 <- c()
# temp_6 <- c()
# temp_ave <- c()
# 
# # Loop through the vector sequence with every third value, then take the average of
# # three consecutive months (i.e. q2)
# for(i in loop_seq) {
#   
#   temp_4<- temp_q2[, , (i)]
#   temp_5 <- temp_q2[, , (i + 1)]
#   temp_6 <- temp_q2[, , (i + 2)]
#   
#   temp_ave <- (temp_4 + temp_5 + temp_6) / 3
#   
#   list_pos <- ((i/3) - (1/3)) + 1 # to get index 1:n(years)
#   
#   dlist[[list_pos]] <- temp_ave
#   
# }
# 
# # Now name the lists with the year:
# names(dlist) <- unique(years_keep)
# 
# # Filter years in the pred-grid data frame to only have the years I have temperature for
# d_sub_temp <- dat %>% filter(year %in% names(dlist)) %>% droplevels()
# 
# # Now keep only the unique combinations of lat lon and year
# d_sub_temp <- d_sub_temp %>% mutate(id_temp = paste(lat, lon, year)) %>% distinct(id_temp, .keep_all = TRUE)
# 
# # Create data holding object
# data_list <- list()
# 
# # Create factor year for indexing the list in the loop
# d_sub_temp$year_f <- as.factor(d_sub_temp$year)
# 
# # Loop through each year and extract raster values for the pred-grid data points
# for(i in unique(d_sub_temp$year_f)) {
#   
#   # Subset a year
#   temp_slice <- dlist[[i]]
#   
#   # Create raster for that year (i)
#   r <- raster(t(temp_slice), xmn = min(lon), xmx = max(lon), ymn = min(lat), ymx = max(lat),
#               crs = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
#   
#   # Flip...
#   r <- flip(r, direction = 'y')
#   
#   plot(r, main = i)
#   
#   # Filter the same year (i) in the pred-grid data and select only coordinates
#   d_slice <- d_sub_temp %>% filter(year_f == i) %>% dplyr::select(lon, lat)
#   
#   # Make into a SpatialPoints object
#   data_sp <- SpatialPoints(d_slice)
#   
#   # Extract raster value (temperature)
#   rasValue <- raster::extract(r, data_sp)
#   
#   # Now we want to plot the results of the raster extractions by plotting the pred-grid
#   # data points over a raster and saving it for each year.
#   # Make the SpatialPoints object into a raster again (for pl)
#   df <- as.data.frame(data_sp)
#   
#   # Add in the raster value in the df holding the coordinates for the pred-grid data
#   d_slice$temp <- rasValue
#   
#   # Add in which year
#   d_slice$year <- i
#   
#   # Create a index for the data last where we store all years (because our loop index
#   # i is not continuous, we can't use it directly)
#   index <- as.numeric(d_slice$year)[1] - 1992
#   
#   # Add each years' data in the list
#   data_list[[index]] <- d_slice
#   
# }
# 
# # Now create a data frame from the list of all annual values
# pred_grid_temp <- dplyr::bind_rows(data_list)
# 
# lims <- pred_grid_temp %>% drop_na(temp) %>% summarise(min = min(temp),
#                                                        max = max(temp))
# 
# # Plot and compare with rasters
# ggplot(pred_grid_temp, aes(lon, lat, color = temp)) +
#   facet_wrap(~year) +
#   scale_colour_gradientn(colours = rev(terrain.colors(10)),
#                          limits = c(lims$min, lims$max)) +
#   geom_sf(data = world, inherit.aes = F, size = 0.2, alpha = 0) +
#   coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
#   geom_point(size = 0.1) +
#   NULL
# 
# # Left join in with the benthic again
# # First make the same selection of years in pred-grid 2
# # sort(unique(dat$year))
# 
# pred_grid_temp <- pred_grid_temp %>% mutate(id_temp = paste(year, lon, lat)) %>% dplyr::select(-year, -lon, -lat)
# 
# dat <- dat %>% mutate(id_temp = paste(year, lon, lat))
# 
# # left_join in temperature
# nrow(dat)
# dat <- left_join(dat, pred_grid_temp)
# nrow(dat)
# 
# ggplot(dat, aes(lon, lat, color = temp)) +
#   facet_wrap(~year) +
#   scale_colour_gradientn(colours = rev(terrain.colors(10)),
#                          limits = c(lims$min, lims$max)) +
#   geom_sf(data = world, inherit.aes = F, size = 0.2, alpha = 0) +
#   coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
#   geom_point(size = 0.1) +
#   NULL
```

Add in ICES areas information

```{r add ICES areas, message=FALSE, results="hide"}
# Get ICES rect
dat$ices_rect <- ices.rect2(lon = dat$Longitud, lat = dat$Latitud)

ggplot(dat, aes(x = Longitud, y = Latitud, color = ices_rect)) +
  geom_point() +
  geom_sf(data = world, inherit.aes = F, size = 0.2) +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_classic(base_size = 20) +
  NULL

# Color according to sub-division
dat <- dat %>%
  mutate(SubDiv = NA) %>%
  mutate(SubDiv = ifelse(ices_rect %in% c("37G0", "37G1",
                                          "38G0", "38G1",
                                          "39F9", "39G0", "39G1",
                                          "40F9", "40G0", "40G1"), "22", SubDiv)) %>%
  mutate(SubDiv = ifelse(ices_rect == "40G2", "23", SubDiv)) %>%
  mutate(SubDiv = ifelse(ices_rect %in% c("37G2", "37G3", "37G4",
                                          "38G1", "38G2", "38G3", "38G4",
                                          "39G1", "39G2", "39G3", "39G4",
                                          "40G1"), "24", SubDiv)) %>%
  mutate(SubDiv = ifelse(ices_rect %in% c("40G4",
                                          "37G5", "37G6", "37G7",
                                          "38G5", "38G6", "38G7",
                                          "39G5", "39G6", "39G7",
                                          "40G5", "40G6", "40G7",
                                          "41G5", "41G6", "41G7"), "25", SubDiv)) %>%
  mutate(SubDiv = ifelse(ices_rect %in% c("37G8", "37G9", "37H0",
                                          "38G8", "38G9", "38H0",
                                          "39G8", "39G9", "39H0",
                                          "40G8", "40G9", "40H0",
                                          "41G8", "41G9", "41H0"), "26", SubDiv)) %>%
  mutate(SubDiv = ifelse(ices_rect %in% c("42G6", "42G7",
                                          "43G6", "43G7",
                                          "44G6", "44G7", "44G8"), "27", SubDiv)) %>%
  mutate(SubDiv = ifelse(ices_rect %in% c("42G8", "42G9", "42H0", "42H1", "42H2",
                                          "43G8", "43G9", "43H0", "43H1", "43H2",
                                          "44G8", "44G9", "44H0", "44H1", "44H2"), "28", SubDiv)) %>%
  mutate(SubDiv = ifelse(Latitud > 58.5 & Longitud > 19, "29", SubDiv)) %>%
  mutate(SubDiv = ifelse(Latitud > 57 & Latitud < 58.5 & Longitud > 19 & Longitud < 22, "28", SubDiv)) %>%
  mutate(SubDiv = ifelse(Latitud > 57 & Latitud < 60 & Longitud > 16 & Longitud < 18, "27", SubDiv)) %>%
  mutate(SubDiv = ifelse(Latitud > 55.5 & Latitud < 56.5 & Longitud > 14 & Longitud < 16, "25", SubDiv)) #%>%
  #mutate(SubDiv = factor(SubDiv))

ggplot(dat, aes(x = Longitud, y = Latitud, color = SubDiv)) +
  geom_point() +
  coord_cartesian(expand = 0) +
  scale_color_brewer(palette = "Dark2") +
  geom_sf(data = world, inherit.aes = F, size = 0.2) +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_classic(base_size = 20) +
  NULL

# Now do a manual area, similar to sub divisions
dat$sub_area <- as.numeric(dat$SubDiv)

unique(dat$sub_area)

# Now modify them...
dat <- dat %>%
  mutate(sub_area = ifelse(Latitud > 56.18 & Longitud < 18, 27, sub_area)) %>% 
  mutate(sub_area = ifelse(Longitud > 17.6 & Latitud >= 56.49, 28, sub_area)) %>% 
  mutate(sub_area = ifelse(Longitud > 17 & Latitud < 56.49, 26, sub_area))

ymin = 55; ymax = 58; xmin = 12.5; xmax = 20

ggplot(dat, aes(x = Longitud, y = Latitud, color = factor(sub_area))) +
  geom_point(size = 0.5) +
  coord_cartesian(expand = 0) +
  geom_sf(data = world, inherit.aes = F, size = 0.2) +
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) + 
  theme_classic(base_size = 20) +
  NULL

ggsave("figures/sub_areas_benthicdata.png", width = 9, height = 9, dpi = 600)
```

Now add in the species 

**Flounder**: Amphipoda, *Saduria entomon*, *Mytilus sp.*, *Limecola
balthica*, "Other inverts"

**Cod**: Polychaeta, Cumacea, *Saduria entomon*, Mysida, "Other inverts"

Let's now make a new category based on the above (I suggest using "Other
inverts" as all remaining taxa):

```{r make new species group, warning=FALSE, message=FALSE}
# First decide how to subset the data to fit the above categories
# Flounder
filter(dat, Ordning == "Amphipoda") # This is easy
filter(dat, taxa == "Saduria entomon") # This is easy
unique(filter(dat, Ordning == "Mytiloida")$taxa) # For Mytilus sp. we can use the "Ordning" Mytiloida
filter(dat, taxa == "Limecola balthica") # This is also easy

# Cod
filter(dat, Klass == "Polychaeta") # Easy
filter(dat, Ordning == "Cumacea") # Easy
filter(dat, Ordning == "Mysida"); filter(dat, Familj == "Mysidae")  # Not sure how to deal with this one, maybe there aren't any in the data. Skip for now

# Ok, based on the species groups, create a new variable for grouping taxa based in importance in cod and flounder diets
dat <- dat %>% 
  mutate(species_group = "Other inverts") %>% 
  mutate(species_group = ifelse(Ordning == "Amphipoda", "Amphipoda", species_group),
         species_group = ifelse(taxa == "Saduria entomon", "Saduria entomon", species_group),
         species_group = ifelse(Ordning == "Mytiloida", "Mytiloida", species_group),
         species_group = ifelse(taxa == "Limecola balthica", "Limecola balthica", species_group),
         species_group = ifelse(Klass == "Polychaeta", "Polychaeta", species_group),
         species_group = ifelse(Ordning == "Cumacea", "Cumacea", species_group)) %>% 
  mutate(species_group = ifelse(is.na(species_group) == TRUE, "Other inverts", species_group)) %>% # Replace the NA with other inverts, because that's where they belong...
  mutate(species_group = factor(species_group))
```

Finally add in depth:
Now add in the depth based on a raster (so that it's identical to the
prediction grid)

```{r add depth, message=FALSE, hide=TRUE}
west <- raster("data/depth_geo_tif/D5_2018_rgb-1.tif")
#plot(west)

east <- raster("data/depth_geo_tif/D6_2018_rgb-1.tif")
#plot(east)

dep_rast <- raster::merge(west, east)

dat$depth_rast <- extract(dep_rast, dat[, 29:28])

# Convert to depth (instead of elevation)
ggplot(dat, aes(depth_rast)) + geom_histogram()
dat$depth_rast <- (dat$depth_rast - max(dat$depth_rast)) *-1
ggplot(dat, aes(depth_rast)) + geom_histogram()

# Compare to built in depth data
ggplot(dat, aes(djup, depth_rast)) + 
  geom_point() +
  geom_abline(color = "red")

ymin = 54; ymax = 58; xmin = 9.5; xmax = 22

dat %>% 
  filter(depth_rast > 0) %>% 
  ggplot(., aes(Longitud, Latitud, color = depth_rast)) + 
  scale_color_viridis() +
  geom_sf(data = world, inherit.aes = F, size = 0.2, fill = NA) +
  geom_point(size = 1) + 
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  NULL
```

We can now save this complete data set to avoid having to run all this
code again

```{r save data, message=FALSE}
# Subset the data a little bit
dat2 <- dat %>%
  rename("depth" = "depth_rast",
         "abundance" = "antal",
         "biomass" = "biomassa",
         "month" = "Month",
         "day" = "Day",
         "year" = "artal",
         "lon" = "Longitud",
         "lat" = "Latitud") %>% 
  dplyr::select(year, month, day, lon, lat, abundance, biomass, species_group, depth, prov_nr, provID, sub_area, taxa)

# Now we can save the data
write.csv(dat2, "data/data_complete.csv")
```

After that, we can also do some visual exploration of the data

```{r explore, fig.align="center", warning=FALSE, message=FALSE}
colnames(dat)

# Color by taxonomic information
# Expand color palette
colourCount = length(unique(dat$Ordning))
getPalette = colorRampPalette(brewer.pal(9, "Set2"))

# Frequency by Stam fill by Klass in data
dat %>%
  ggplot(., aes(x = forcats::fct_infreq(Stam), fill = Klass)) +
  geom_bar(color = "grey1", size = 0.1) +
  coord_cartesian(expand = 0) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = getPalette(colourCount)) +
  ggtitle("Occurence in data Stam/Klass") +
  xlab("Stam") +
  NULL

# Frequency by Klass fill by Ordning in data
dat %>%
  ggplot(., aes(x = forcats::fct_infreq(Klass), fill = Ordning)) +
  geom_bar(color = "grey1", size = 0.1) +
  coord_cartesian(expand = 0) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = getPalette(colourCount)) +
  ggtitle("Occurence in data Klass/Ordning") +
  xlab("Klass") +
  NULL

# Filter the most common "Klass" and plot by year
common_class <- 
  c("Bivalvia", "Gastropoda", "Insecta", "Malacostraca", " Polychaeta", "Enopla")

# We'll split this plot into 2 to get a better view

# Plot 1/2
dat %>%
  filter(Klass %in% common_class & artal < 1999) %>% 
  ggplot(., aes(x = forcats::fct_infreq(Klass), fill = Ordning)) +
  geom_bar(color = "grey1", size = 0.1) +
  facet_wrap(~artal) +
  coord_cartesian(expand = 0) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = getPalette(colourCount)) +
  ggtitle("Occurence in data Klass/Ordning") +
  xlab("Klass") +
  NULL

# Plot 2/2
dat %>%
  filter(Klass %in% common_class & artal > 1999) %>% 
  ggplot(., aes(x = forcats::fct_infreq(Klass), fill = Ordning)) +
  geom_bar(color = "grey1", size = 0.1) +
  facet_wrap(~artal) +
  coord_cartesian(expand = 0) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = getPalette(colourCount)) +
  ggtitle("Occurence in data Klass/Ordning") +
  xlab("Klass") +
  NULL

# Biomass by Klass fill by Ordning in data
dat %>%
  drop_na("biomassa") %>% 
  group_by(Klass, Ordning) %>% 
  summarise(tot_bio = sum(biomassa)) %>% 
  ggplot(., aes(x = reorder(Klass, -tot_bio, sum), y = tot_bio, fill = Ordning)) +
  geom_bar(stat = "identity") +
  coord_cartesian(expand = 0) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = getPalette(colourCount)) +
  ggtitle("Biomass Klass/Ordning") +
  xlab("Klass") +
  NULL

# Abundance by Klass fill by Ordning in data
dat %>%
  drop_na("antal") %>% 
  group_by(Klass, Ordning) %>% 
  summarise(tot_abund = sum(antal)) %>% 
  ggplot(., aes(x = reorder(Klass, -tot_abund, sum), y = tot_abund, fill = Ordning)) +
  geom_bar(stat = "identity") +
  coord_cartesian(expand = 0) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_manual(values = getPalette(colourCount)) +
  ggtitle("Abundance Klass/Ordning") +
  xlab("Klass") +
  NULL

#... And many more plots can be made like this, just change filters and variables
```

We can also plot the data on maps...

```{r plot maps, warning=FALSE, message=FALSE}
# Plot sample intensity in space and time
dat %>%
  filter(Klass %in% common_class) %>%
  distinct(provID, .keep_all = T) %>%
  ggplot(., aes(x = Longitud, y = Latitud)) +
    geom_point(size = 0.1, color = "red") +
    facet_wrap(~artal, ncol = 10) +
    coord_cartesian(expand = 0) +
    geom_sf(data = world, inherit.aes = F, size = 0.2) +
    coord_sf(xlim = xlim, ylim = ylim) + 
    ggtitle("Locations of samples (all)") +
    theme_map() +
    theme(axis.text.x = element_text(angle = 90)) +
    NULL
```

We can also plot time series by station, and for species groups used in
Haase *et al.* (2020) for describing the diet of cod and flounder

More plots...

```{r more sample plots, message=FALSE}
# Plot the annual average of the total biomass of the species_group within a sample
dat %>% 
  group_by(species_group, artal, provID) %>% 
  drop_na(biomassa) %>% 
  summarise(tot_biomass_sample = sum(biomassa)) %>% 
  ungroup() %>% 
  group_by(species_group, artal) %>% # Now average these by year and by species
  mutate(ave_biomass = mean(tot_biomass_sample)) %>% # Calculate mean of total biomass by species group per unique sample
  ungroup() %>% 
  ggplot(., aes(artal, ave_biomass, fill = species_group)) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Set2")

# Same as above but with abundance
dat %>% 
  group_by(species_group, artal, provID) %>% 
  drop_na(antal) %>% 
  summarise(tot_abun_sample = sum(antal)) %>% 
  ungroup() %>% 
  group_by(species_group, artal) %>%
  mutate(ave_abun = mean(tot_abun_sample)) %>%
  ungroup() %>% 
  ggplot(., aes(artal, ave_abun, fill = species_group)) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Set2")

# Looks like something happened in ~1980 - likely due to spatial extent of sampling increasing introducing a bias

# Plot normalized biomasses on a map from 2015 and forward (from when we have data on both cod and flounder diets)
dat %>%
  filter(!species_group == "Other inverts") %>% 
  group_by(species_group, artal, Latitud, Longitud) %>% 
  drop_na(biomassa) %>% 
  summarise(tot_biomass_sample = sum(biomassa)) %>% 
  ungroup() %>% 
  filter(artal > 2014) %>% 
  filter(species_group %in% unique(species_group)[1:3]) %>% 
  group_by(species_group) %>% 
  mutate(max_tot_biomass = max(tot_biomass_sample)) %>%
  ungroup() %>% 
  mutate(tot_biomass_sample_norm = tot_biomass_sample / max_tot_biomass) %>% 
  ggplot(., aes(x = Longitud, y = Latitud, color = log(tot_biomass_sample_norm))) +
    geom_point() +
    facet_grid(species_group ~ artal) +
    coord_cartesian(expand = 0) +
    geom_sf(data = world, inherit.aes = F, size = 0.2) +
    coord_sf(xlim = xlim, ylim = ylim) + 
    ggtitle("Biomass normalized to max biomass by species group and location") +
    theme_map() +
    theme(axis.text.x = element_text(angle = 90)) +
    scale_color_viridis() +
    NULL

# Same plot for the remaining species groups
dat %>%
  filter(!species_group == "Other inverts") %>% 
  group_by(species_group, artal, Latitud, Longitud) %>% 
  drop_na(biomassa) %>% 
  summarise(tot_biomass_sample = sum(biomassa)) %>% 
  ungroup() %>% 
  filter(artal > 2014) %>% 
  filter(species_group %in% unique(species_group)[4:6]) %>% 
  group_by(species_group) %>% 
  mutate(max_tot_biomass = max(tot_biomass_sample)) %>%
  ungroup() %>% 
  mutate(tot_biomass_sample_norm = tot_biomass_sample / max_tot_biomass) %>% 
  ggplot(., aes(x = Longitud, y = Latitud, color = log(tot_biomass_sample_norm))) +
    geom_point() +
    facet_grid(species_group ~ artal) +
    coord_cartesian(expand = 0) +
    geom_sf(data = world, inherit.aes = F, size = 0.2) +
    coord_sf(xlim = xlim, ylim = ylim) + 
    ggtitle("Biomass normalized to max biomass by species group and location") +
    theme_map() +
    theme(axis.text.x = element_text(angle = 90)) +
    scale_color_viridis() +
    NULL

# Plot samples of saduria for all years
dat %>%
  filter(species_group == "Saduria entomon") %>%
  group_by(artal, Latitud, Longitud) %>% 
  drop_na(biomassa) %>% 
  summarise(tot_biomass_sample = sum(biomassa)) %>% 
  ungroup() %>% 
  mutate(max_tot_biomass = max(tot_biomass_sample)) %>%
  ungroup() %>% 
  mutate(tot_biomass_sample_norm = tot_biomass_sample / max_tot_biomass) %>% 
  ggplot(., aes(x = Longitud, y = Latitud, color = log(tot_biomass_sample_norm))) +
    geom_point() +
    facet_wrap(~ artal, ncol = 10) +
    coord_cartesian(expand = 0) +
    geom_sf(data = world, inherit.aes = F, size = 0.2) +
    coord_sf(xlim = xlim, ylim = ylim) + 
    ggtitle("Biomass normalized to max biomass across samples") +
    theme_map() +
    theme(axis.text.x = element_text(angle = 90)) +
    scale_color_viridis() +
    NULL

# Plot time series of the annual average of the total biomass in each sample by species_group. This assumes sampling locations haven't changed in a way leading to bias.
# Filter years after 1983 because of the low sample size before that
dat %>% 
  filter(!species_group == "Other inverts") %>% 
  group_by(species_group, artal, provID) %>% 
  drop_na(biomassa) %>% 
  summarise(tot_biomass_sample = sum(biomassa)) %>% 
  ungroup() %>% 
  group_by(species_group, artal) %>% # Now average these by year and by species group
  mutate(ave_biomass = mean(tot_biomass_sample)) %>% # Calculate the annual mean of total biomass by species group per unique sample
  ungroup() %>%
  filter(artal > 1983) %>% 
  ggplot(., aes(x = artal, y = ave_biomass, color = species_group)) +
    geom_point(size = 2) +
    stat_smooth(formula = y ~ s(x, bs = "cs", k = 3)) +
    facet_wrap(~ species_group, scales = "free_y") +
    coord_cartesian(expand = 0) +
    theme(aspect.ratio = 1, legend.position = "bottom") +
    scale_color_brewer(palette = "Set2") +
    NULL
```

Interesting!

However, a proper index should be calculated to account for the
spatiotemporal correlation structure of the residuals, and thus also the
changes in sampling stations across years. And for the compound groups
in particular, we should make sure that we measure the same species each
year.

So, which selections and group structure all boils down to the tradeoff
between spatial, temporal and taxonomic resolution.

-   **Environmental variables**: Can be added either by using model
    outputs from SMHI (available between 1993-2019) or by selecting
    specific stations that are representative of the area (e.g. depth
    stratified)

-   **Stomach data & individual-level biological data**: Adding newer
    data is in progress - waiting for samples.

-   **Catch data (for index of abundance or density)**: Latest data can
    be downloaded from DATRAS and matched to stomach data
